---
title: "tø uge 2"
output:
  pdf_document: default
  html_document: default
date: "2025-01-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(resampledata3)
```


Compute the mean x and median m of the eight numbers $1, 2, 4, 5, 6, 8, 11$
15. Let $$f (x) = \sqrt{x}$$.\newline

Apply this transformation to the data, and then com- pute the mean $x̃$ and the median $m̃$ of the transformed data

question a: 
$$\text{ Is}\,\, f (x) = x̃?$$\newline
question b: 
$$\text{Is}  f ( m ) = m̃?$$

```{r}
#we write up the data 
x=c(1,2,4,5,6,8,11)
#calculate the transformed data
fx=sqrt(x)
# find the mean for the data
xb=mean(x)
#then the mean of the transformation 
mean(fx)
#then f of the mean 
sqrt(xb)
#now we do the exactly the same but for the median 
m=median(x)
median(fx)
sqrt(m)
```

the mean is not equal since f is concave then  
$$f(\dfrac{1}{n}*\sum_{i=1}^nx_i)\leq \dfrac{1}{n} \sum_{i=1}^nf(x_i)$$ 
the median in this case is equal but this is only true in general when f in linear or n is uneven and f is montone




Let x and m denote the mean and median, respectively, of $$x_1 < x_2 < · · · < x_n$$ 
Let f be a real-valued function.

a) Is f(x) the mean off $$f(x_1),f(x_2),...,f(x_n)$$?
no simply look at the last exercise 


b) Is f(m) the median off f(x1),f(x2),...,f(xn)?
try $x=(x_1,x_2)=(1,2)$
$f=x^2$
not equal
c) Are there any conditions that would ensure that f (x) is the mean of
the transformed data?
liniarity of f or no variance in x


d) Are there any conditions that would ensure that f (m) is the median
of the transformed data?

liniarity of f or no variance in x 
or f order preserving and n uneven


```{r}
s=table(FlightDelays$DepartTime)
barplot(s)
```

```{r}
tab <- table(FlightDelays$Day,FlightDelays$Delayed30)
prop.table(tab,1)
```

```{r}
boxplot(FlightDelays$FlightLength~FlightDelays$Delayed30)

```


opgave 2.4,d
vi ser at hvis flyet er forsiket er der en støøre chance for at det er en medium længde flyvetur men også mange flere outliers, jeg ville kke drage nogle konklutioner


```{r}
par(mfrow=c(1,2))
tab <- table(Recidivism$Recid)
barplot(tab)
prop.table(tab)
tab2 <-table(Recidivism$Age25,Recidivism$Recid)
prop.table(tab2,1)
```


wee seee that about 30,6% of offenders over 25 go back to prison while about 36.5% of under 25 offenders go back 



```{r}
t=Recidivism$Days
s=Recidivism$Type
boxplot(t~s)

```


the median reeviction time is lover for tech crimes . all quantiles are lover for the tech crime type

```{r}
quantile(Recidivism$Days,na.rm=TRUE)
```

```{r}

par(mfrow=c(1,2))
over25 <-subset(Recidivism, Age25== "Over 25")
under25 <- subset(Recidivism,Age25== "Under 25")

plot.ecdf(over25$Days)
plot.ecdf(under25$Days)
```

opgave 1.26

```{r}
curve(log(x),0,2,10000)
abline(h=0)
abline(v=1)

```

```{r}
curve(x*sin(x^2),0,2*pi,1000)
curve(x*cos(x^2),0,2*pi,1000,add = TRUE)

```

opgave 1.27

```{r}
data <-TXBirths2004
hist(TXBirths2004$Weight, freq= FALSE, breaks=250*0:20)


curve(dnorm(x,mean=mean(TXBirths2004$Weight),sd=sd(TXBirths2004$Weight)),0,5000, add=TRUE)
```

```{r}
hist(TXBirths2004$Weight, freq= FALSE, breaks=50*0:100)
curve(dnorm(x,mean=mean(TXBirths2004$Weight),sd=sd(TXBirths2004$Weight)),0,5000, add=TRUE)
```


histogrammet bliver langt mindre overskueligt og mere flimret 
og derved sværere at konkludere ud fra 
}


oms opgave 1.25
part a)



$$\hat{F}_n(x)=\dfrac{1}{n}*\sum_{i=1}^n\mathbb{1}_{]-\infty,x]}(X_i)$$



vi noterer at:  $$\mathbb{1}_{]-\infty,x]}(X_i)=1\Leftrightarrow X_i\leq x$$ 
og $$\mathbb{1}_{]-\infty,x]}(X_i)=0\Leftrightarrow X_i> x$$ 
da er denne transformation bernulli med sandylingheds paramete
$$\mathbb{P}(X_i \leq x)=F(x)$$
 summen af i.i.d bernulli er binomial
$$n*\hat{F}_n(x)=\sum_{i=1}^n\mathbb{1}_{]-\infty,x]}(X_i)$$  er da binomial

del b)


$$\mathbb{E}[\hat{F}_n(x)]=\dfrac{1}{n}*\mathbb{E}[n*\hat{F}_n(x)]=F(x)$$


$$\mathbb{V}[\hat{F}_n(x)]=\dfrac{1}{n^2}*\mathbb{V}[n*\hat{F}_n(x)]=\dfrac{1}{n^2}*n*F(x)*[1-F(x)]$$

$$=\dfrac{1}{n}*F(x)*(1-F(x))$$

del c) vis

$$\lim_{n\to \infty}\mathbb{P}(\lvert \hat{F}_n(x)-F(x)\lvert \geq \delta)=0 \quad \forall  \delta>0$$


vi bemærker at  $$\hat{F}_n(x)=\dfrac{1}{n}*\sum_{i=1}^n\mathbb{1}_{]-\infty,x]}(X_i)$$
er en stikprøvemiddelværdi for varablen vi allerede har beskrevet som bernulli
da siger store tals lov
$$ \overline X \to^{\mathbb{P}} \mu \quad for\,\, n\to\infty $$

da  $$\hat{F}_n(x)$$  er en stkprøve middelværdi for bernulli må:
 $$\hat{F}_n(x) \to^{\mathbb{P}}F(x)  \quad for\,\, n\to\infty $$






